### 2013-08-02 Doc/MMVR2014/abstract.tex
FIX: finalization of the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@608 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{frontmatter}%
\begin{abstract}%
Accurate biomechanical modeling of liver is of a paramount interest for pre-operative planning or computer-aided per-operative guidance. 
However, such simulations remain a challenging task, since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, 
each having different mechanical properties. 
%
%While a real-time simulation of vascularized liver has been already described in~\cite{Peterlik2012}, the Glisson's capsule has not been included in the modeling. 
%In~\cite{Hollenstein2006,Ahn2010} it is shown that the capsule plays an important role mainly in the vicinity of the surgical tools. 
%Therefore, the accuracy of the liver response in the vicinity of the surgical tool requires correct modeling of the capsule as a component of the liver model. 
%
In this paper we propose a complete liver model, where the parenchyma is modeled as a finite element volume, the vessels are modeled as series of beam elements, and the   corotational membrane elements based on constant-strain formulation, which are coupled mechanically to the underlying tetrahedra. 
We show that using this model we are able to reproduce accurately biomechanical tests, while maintaining the real-time aspect of the simulation.

>>>>

\begin{frontmatter}%
\begin{abstract}%
Accurate biomechanical modeling of liver is of a paramount interest for pre-operative planning or computer-aided per-operative guidance. 
However, such simulations remain a challenging task, since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, 
each having different mechanical properties. 
%
%While a real-time simulation of vascularized liver has been already described in~\cite{Peterlik2012}, the Glisson's capsule has not been included in the modeling. 
%In~\cite{Hollenstein2006,Ahn2010} it is shown that the capsule plays an important role mainly in the vicinity of the surgical tools. 
%Therefore, the accuracy of the liver response in the vicinity of the surgical tool requires correct modeling of the capsule as a component of the liver model. 
%
In this paper we propose a complete liver model, where the parenchyma is modeled as a finite element volume, the vessels are modeled as series of beam elements, 
and the corotational membrane elements based on constant-strain formulation are used for capsule. These components are coupled mechanically as shown in the paper.
We show that using this model we are able to reproduce accurately biomechanical tests, while maintaining the real-time aspect of the simulation.

### 2013-08-02 Doc/MMVR2014/intro.tex
FIX: finalization of the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@608 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

Computer-aided physics-based medical simulation has proven to be an extremely useful technique in the area of medical training. 
Nevertheless, whereas generic models are usually required in training simulators, accurate patient-specific modeling
becomes necessary as soon as computer simulation is to be employed in the pre-operative planning. 
%At the same time, 
%interactivity of such models remains an important aspect, requiring real-time simulation which is often difficult to 
%achieve given the complexity of soft tissues. 
%
When simulating the behavior of human liver, the task of real-time accurate modeling is challenging, mainly because of the complex structure 
of this organ composed of three main constituents: \emph{parenchyma}, \emph{vascular networks} and \emph{Glisson's capsule}.
The parenchyma has certainly been the most studied component of the liver; actually researchers agree on hyperelastic 
properties of the tissue, for which the mechanical parameters have been reported for example in~\cite{Kerdok2006,Gao2009}. 
Moreover, several methods have been proposed to model the hyperelastic behaviour at real-time rates, such as multiplicative Jacobian decomposition
introduced in~\cite{Marchesseau2010}.
%
%\TG{The following paragraph is optional. Unnecessary but adds to the global context}
The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that the 
influence of vessels on the mechanical behaviour of the organ is significant. %, mainly if large deformations occur. 
%Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
%the authors propose a composite model allowing for real-time simulation of entire liver with vascularization.
Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall), 
the authors propose a composite model allowing for real-time simulation.

@@

In this paper we present preliminary results of our work on the complete liver model. 
We present a real-time composite model accounting for parenchyma and Glisson's capsule, compatible with previously proposed vascularized model~\cite{Peterlik2012}.
%The model is based on two different finite element representations for each constituent coupled together.
%The main contribution of the paper is twofold: first, we present a real-time model of the liver, including
%the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
%linked together via mechanical coupling.
%We show that the model mimics the \emph{local} experiments described in~\cite{Hollenstein2006}.

>>>>

Computer-aided physics-based medical simulation has proven to be an extremely useful technique in the area of medical training. 
Whereas generic models are usually required in training simulators, accurate patient-specific modeling
becomes necessary as soon as computer simulation is to be employed in the pre-operative planning. 
%At the same time, 
%interactivity of such models remains an important aspect, requiring real-time simulation which is often difficult to 
%achieve given the complexity of soft tissues. 
When simulating the behavior of human liver, the task of real-time accurate modeling is challenging, mainly because of the complex structure 
of this organ composed of three main constituents: \emph{parenchyma}, \emph{vascular networks} and \emph{Glisson's capsule}.
The parenchyma has certainly been the most studied component of the liver; actually researchers agree on hyperelastic 
properties of the tissue, for which the mechanical parameters have been reported for example in~\cite{Kerdok2006}.%,Gao2009}. 
Moreover, several methods have been proposed to model the hyperelastic behaviour at real-time rates, such as multiplicative Jacobian decomposition
introduced in~\cite{Marchesseau2010}.
%
%\TG{The following paragraph is optional. Unnecessary but adds to the global context}
The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that the 
influence of vessels on the mechanical behaviour of the organ is significant. %, mainly if large deformations occur. 
%Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
%the authors propose a composite model allowing for real-time simulation of entire liver with vascularization.
Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall), 
the authors propose a composite model allowing for real-time simulation.

@@

In this paper we present preliminary results of our work on the complete liver model. 
We present a real-time composite model accounting for parenchyma and Glisson's capsule, compatible with previously proposed vascularized model~\cite{Peterlik2012}.
We also show that the capsule has a significant impact on the mechanical response of the organ. 
%The model is based on two different finite element representations for each constituent coupled together.
%The main contribution of the paper is twofold: first, we present a real-time model of the liver, including
%the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
%linked together via mechanical coupling.
%We show that the model mimics the \emph{local} experiments described in~\cite{Hollenstein2006}.

### 2013-08-02 Doc/MMVR2014/results.tex
FIX: finalization of the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@608 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

In Fig.~\ref{fig-aspiration2} the profiles of cuts in the middle of the
test cube are presented showing significantly larger deformation in the model without capsule assuming that the same negative pressure was applied 
on the tissue surface inside the tube. Moreover, it can be observed that the deformation without capsule is overestimated by a factor 
close to 2, which is in perfect agreement with the results reported in~\cite{Hollenstein2006}.
Fig.~\ref{fig-aspiration3} presents the evolution of displacement in time at central node on the cube surface. The effect of increased combined stiffness is clearly visible.

Further, we employed the simulation to reproduce the displacements values obtained via simulation in~\cite{Hollenstein2006,Nava2008}.
Whereas the simulation in the referenced work was done on a mesh composed of tetrahedral elements for both the parenchyma 
and the capsule, we modeled the capsule using the triangular elements as shown in section~\ref{ss:capsuleModel}.

@@

\subsection{Global Deformation of Complete Model}
The complete model was constructed using contrast-enhanced CT data acquired on a female pig
being positioned first in supine and then flank position.
The liver and hepatic vein were segmented for both datasets  using semi-automatic methods available in ITKSnap~\footnote{www.itksnap.org}.
A tetrahedral mesh was obtained from the segmented maps using CGAL~\footnote{www.cgal.org}.
The semi-automatic method presented in~\cite{Peterlik2012} was
then applied to the segmented map of the hepatic vein, resulting in linked-beam model.
Finally, the model of the Glisson's capsule was constructed using the surface elements extracted from the volume mesh.
%described in section~\ref{ss:capsuleModel} over the surface triangles extracted from the tetrahedral mesh.
The complete model composed of 4266 tetrahedra, 188 beams and 1284 triangles was 
used in simulation where external force representing the gravity was applied to the model,
resulting in large deformations mainly in the area of lobes. The refresh rate of 22\,FPS
was achieved on a desktop computer with Intel i7 CPU running at 3.4\,GHz and 8\,GB of memory.
The simulation with gravity was also used as a first demonstration of the global influence of the capsule.
In Fig.~\ref{f:global}a, the models with capsule (red) and without (yellow) are depicted after being
deformed due to the gravity, showing a significant difference between the two models.

>>>>

In Fig.~\ref{fig-aspiration2} the profiles of cuts in the middle of the
test cube are presented showing significantly larger deformation in the model without capsule assuming that the same negative pressure was applied 
on the tissue surface inside the tube. Moreover, it can be observed that the deformation without capsule is overestimated by a factor 
close to 2, which is in perfect agreement with the results reported in~\cite{Hollenstein2006}.
Fig.~\ref{fig-aspiration3} presents the evolution of displacement in time at central node on the cube surface showing clear the effect of increased combined stiffness.

Further, we employed the simulation to reproduce the displacements values obtained via simulation in~\cite{Hollenstein2006,Nava2008}.
Whereas the simulation in the referenced work was done on a mesh composed of tetrahedral elements for both the parenchyma 
and the capsule, we employ the membrane CST elements.

@@

\subsection{Global Deformation of Complete Model}
The complete model was constructed using contrast-enhanced CT data acquired on a female pig.
The liver and hepatic vein were segmented using semi-automatic methods available in ITKSnap~\footnote{www.itksnap.org}.
A tetrahedral mesh was obtained from the segmented maps using CGAL~\footnote{www.cgal.org}.
The semi-automatic method presented in~\cite{Peterlik2012} was
then applied to the segmented map of the hepatic vein, resulting in linked-beam model.
Finally, the model of the Glisson's capsule was constructed using the surface elements extracted from the volume mesh.
%described in section~\ref{ss:capsuleModel} over the surface triangles extracted from the tetrahedral mesh.
The complete model composed of 4266 tetrahedra, 188 beams and 1284 triangles was 
used in simulation where external force representing the gravity was applied to the model,
resulting in large deformations mainly in the area of lobes. The refresh rate of 22\,FPS
was achieved on a desktop computer with Intel i7 CPU running at 3.4\,GHz and 8\,GB of memory.
The simulation with gravity was also used as a first demonstration of the global influence of the capsule.
In Fig.~\ref{f:gravity}a, the models with capsule (red) and without (yellow) are depicted after being
deformed due to the gravity, showing a significant difference between the two models.
